Application of Micro-Rotors to Induce Crystal Nucleation: Challenges, Takeaways, and Future Prospects\\ PhD Thesis

### 2025-06-27 Latex_files/Thesis/chapters/chapter 3/main.tex
27/06/2025 chapters 1,3,4 and structural changes

\chapter{Effects of localised shearing on crystal growth and nucleation}
As outlined in Chapter 1, part of the projects aim is to investigate 
the possibility of using optical tweezing to induce nucleation by 
generating fluid flow within a supersaturated solution. The intent 
of which would be twofold: Firstly to have a repeatable means of 
inducing nucleation under different solution conditions. And secondly, 
to understand the influence of shearing on nucleation at a micro 
level as compared to results in bulk fluid.

It has been shown that for macro-scale systems, the likelihood of 
nucleation increases to a maximum value under increased shearing 
\cite{Debuysschere2023, Mura2016}. Mura and Zaccone developed a 
theoretical framework to describe how the a newly formed nucleus 
experiences two additional growth factors when placed in a moving 
fluid. Firstly, due to increased molecular transport of solute 
molecules the nucleation rate is enhanced in low to moderate fluid 
flows. But in addition, due to shear flow the crystal surface 
undergoes deformation which suppresses the nucleation rate 
undergoing faster fluid flow \cite{Mura2016}. Experimental results 
with glycine solution support this theory; Debuysschere \textit{et 
al} demonstrated that the nucleation rate of supersaturated glycine 
was enhanced up until $\dot{\gamma}\approx3000\ s^{-1}$ 
\cite{Debuysschere2023}. After which the nucleation rate began to 
decrease but was still greater compared to the case where fluid 
flow was minimal.

Optical tweezers can been used to rotate a whole host of microscopic 
objects, with the fastest reported results exceeding $1000\ Hz$ in 
heavy water \cite{Arita2016}. If such a micro-rotor was suspended in
a supersaturated solution the fluid flow around it could be fast enough
that the nucleation rate is locally enhanced. We focused on two 
primary candidates for rotation, Vaterite and 
4-Heptyl-4-biphenylcarbonitrile (7CB). The former being a polymorph of 
calcium carbonate and the latter an example of nematic liquid crystals, 
both of which have been used repeatedly in previous micro-rotor research
\cite{Saito2022, Arita2016, Parkin2009}. In addition, we also consider 
the application of using techniques beam steering to generate fluid 
flow by trapping silica micro-beads. In this instance the fluid flow is 
generated not due to the transfer of angular momentum, but due to 
shearing caused by a moving sphere through stagnant fluid.

To begin with, the discussion of the necessary optical equipment is 
covered, drawing attention to specialised components and techniques 
that are not standard in optical trapping set ups.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
\section{Optical Tweezer Equipment}
In general, all optical tweezers require a laser driver, a focusing 
microscope objectives, a position controller, and position detector 
\cite{Gieseler2020}. The laser used for this project was a $1064\ 
nm$ near infrared laser - provided by CNI Lasers – that was focused 
by a Nikon 100x oil immersion lens. The choice of an oil immersion 
lens is important as the optical oil used prevents a loss of focus 
when used on a glass cover slip. Now, experimental work has shown 
that the trapping efficiency increases with beam diameter up until 
it exceeds $\frac{2}{3}D_{obj}$ \cite{kim2003dependence} where 
$D_{obj}$ is the diameter of the objective aperture. To expand the 
beam front we utilise a Galilean beam expansion arrangement 
(indicated by $f_1$, and $f_2$ in Fig.~\ref{fig:setup}) as recommended 
for high power laser applications. In our initial experiments the 
beam expansion provides a $4\times$ magnification. Whereas in later 
experiments we utilised a galvano-mirror the beam expansion is 
$3\times$ and then the 4f correlator (see \ref{sec:4f_correlation})
provides a further $1.25\times$ magnification (using $f_3$ and $f_4$) 
- the magnification is given by.
\begin{align}
	\frac{D_2}{D_1} = \frac{f_2}{f_1}
\end{align}

It should be noted that the galvano-mirror requires the use of a 
Keplerian beam expansion arrangement which reduces the transmitted 
laser power due to localised heating of the air. Afterwards the 
laser is passed through a dichroic mirror that separates incoming 
infrared and visible light, this is to prevent the laser from 
damaging the CCD camera used for imaging the trapping plane. The 
laser is then focused to a diffraction limited spot by the objective. 
Utilizing a high numerical aperture objective enhances the gradient 
force at the focal point; the trade-off being that the for higher NA 
objectives the trapping depth is reduced due to spherical aberrations. 
While it is possible to increase the trapping depth \cite{Reihani2006} 
by adjusting the objective's tube length this approach is incompatible 
with our trapping arrangement. A 0.25 NA condenser objective refocuses 
the scattered laser light and also provide an aperture for an imaging 
LED to illuminate the focal plane. Samples are loaded onto a piezo 
driven table to that is inserted between the trapping and condensing 
objectives; the piezo drivers allow for sub-micron control of the 
beam focus position to a degree as small as a 10 nm. To detect and 
monitor the position of a trapped particle a quadrant photo diode 
(QPD) was utilised. 
\begin{figure}[h!]
	\centering
	\includegraphics[height=\linewidth, angle=270]{tweezer_setup.pdf}
	\caption{Optical tweezer set up used for the majority of the PhD. 
		The focal lengths of $f_1$, $f_2$, $f_3$, \& $f_4$ are $-20\ mm$, 
		$60\ mm$, $150\ mm$, \& $200\ mm$ respectively [M = mirrors,
		DM = dichroic mirrors, f = focal lenses]. Diagram not 
		drawn to scale.}
	\label{fig:setup}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Position detection methods}
In order to accurately capture the dynamics of a trapped particle, a 
position detection system is required. There are 3 possible methods 
of position detection: video-analysis, lateral-effect position sensing, 
and photodiodes.

Video analysis is ideally suited for multiple traps or situations where 
precision is not the top priority. Whereas lateral-effect and photodiode 
position detection are two examples of back-focal plane interferometry, 
where the interference pattern produced by the target particle is 
extrapolated to determine its position. In order for video analysis to 
match the accuracy of back-focal plane interferometry methods requires 
the camera's frame rate to exceed $1\ kHz$ which can be difficult to 
achieve while maintaining a decent resolution \cite{Gibson2008}. In 
comparison off the shelf back-focal plane detectors can achieve temporal 
resolutions anywhere from $10-100\ kHz$ \cite{BergSoerensen2004}.

A quadrant photo diode (QPD) is a frequently used position detection 
system for optical tweezers due to their high sampling rate, high 
degree of precision, and ease of set up. The QPD is constructed of 
four photo diodes assembled in a quadrant formation, when a particle 
is trapped the interference pattern produced is focused onto the QPD, 
with the maximum intensity mapping to the particle's centre of mass. 
By summing the voltages of the horizontal and vertical quadrants together 
the particle's centre of mass is tracked in the x-y plane. Axial 
displacement can be estimated by observing the change in the total 
voltage of the QPD. The outputted signal gives an indication of the 
particle's relative displacement from the beam focus, but in order to 
convert the signal to distance units the trap needs to be calibrated 
(assuming a linear response curve).
\begin{figure}[h!]
	\centering
	\includegraphics[height=\linewidth, angle=270]{QPD_Lateral_effect.pdf}
	\caption{Comparison between QPD and Lateral effect photodiodes.
	The four quadrants of a QPD (left) experience different photocurrents
	based on the total intensity of light incident on each section 
	(labelled A, B, C, D). 
	Whereas a Lateral effect sensor (right) uses the resistive properties
	of the photodiode surface to vary the create different photocurrents 
	passing through the anodes A, B, C, and D.}
\end{figure}

A lateral-effect sensor has a similar output but works using a the 
entire sensor as a single cell analogous to the focal plane of the 
trapping beam. The four corners of the sensor act as anodes connected 
to a base plate cathode, as the beam moves across the surface of the 
detector each anode will experience a different photocurrent depending 
on how close the centre of the interference pattern is to each anode. 
The advantage of a lateral effect detector is that the linear regime 
is much larger than a QPD making it much better for monitoring the 
position of a trapped particle. However, Lateral-effect sensors are 
often limited in their spacial resolution due to high signal-to-noise 
ratios, requiring a high intensity of light on the sensor in order to 
get a clean signal. As a result, most optical force measurements are 
conducted using a QPD as opposed to a lateral-effect sensor, as often 
the displacement is small enough that the signal-displacement curve 
can be considered linear.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Fourier Optics and 4f correlators}
\label{sec:4f_correlation}
As shown in \ref{fig:setup}, after the Galvano mirror there are a pair
of focal lenses ($f_3$ and $f_4$) that do not seem to serve a clear
purpose. These are in fact crucial for the operation of the Galvano, the
pair of them can be more accurately called a 4f correlator.

A 4f correlator is an example of Fourier optics in practice, understanding
that a focused lens takes a Fourier transform of the light profile. 
Consider a laser with a circular Gaussian profile, if you were to place 
a detector there you would pick up the intensity as a function of its 
position within the beam. If however you focused the light into a single 
point (using a +ve focal lens) you are actually seeing a measurement of 
the phase of your laser with position, in which you would see a diffraction
limited spot ($d = \lambda/2nsin(\theta)$), indicating that the laser is 
collimated. In imaging systems, a series of focal lenses can be used to 
filter out unwanted scattering from an image (or in an inverse case 
differentiate between different images), the placement of each lens is 
shown via fig.~\ref{fig:signal_generator}.
\begin{figure}[h!]
	\centering
	\begin{subfigure}{0.475\linewidth}
		\includegraphics[width=\linewidth]{signal_generator.jpg}
		\caption{}
	\end{subfigure}
	\begin{subfigure}{0.475\linewidth}
		\includegraphics[width=\linewidth, height=0.7\linewidth]
		{galvano_mirror_controllers.jpg}
		\caption{}
	\end{subfigure}

	\caption{Signal generator galvano mirror controller, channel 1 controls the x-axis 
		mirror, while channel 2 controls the y-axis mirror. Both channels can be manipulated
		independently.}
	\label{fig:signal_generator}
\end{figure}

For our applications a 4f correlator is utilised to ensure that the motion
of the galvano-mirrors does not move the focal point of the laser, allowing 
for a stable trap even while in motion. As shown in Fig.~\ref{fig:setup}, after
the galvano-mirror we have our two lenses - $f_3$ and $f_4$ - the former being 
installed $150\ mm$ from the second mirror of the galvano, and the latter being 
installed $200\ mm$ from the back focal plane of the trapping objective. The 
signal generator used was supplied by 'MCS Test Equipment Ltd', allowing for 
dual channel signal control. This allowed us to precisely control the alignment,
amplitude, phase, and frequency of both mirrors making alignment much easier.

While in theory the frequency can be increased until the mechanical limit of the 
mirrors is reached ($\approx 1000\ Hz$), the practical upper limit is determined 
by the trap strength. For a silica sphere this is on the order of $300\ Hz$ when
suspended in water. Likewise while in theory the signal amplitude can be increased
until the voltage limit of the motors is reached ($\approx 5.0 V$) the geometry 
of the focal lenses limits the maximum amplitude to $0.5 V$ as any greater will
move the laser beyond the lens. For basic trapping calibration the galvano-mirrors 
were set to DC output, providing a fixed spot which operates like a typical optical 
trap.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Synthesis of Birefringent Micro spheres}
\label{sec:vaterite}
There are several options for particles that can be rotated 
using optical tweezers \cite{Parkin2009, Saito2022}. Over 
the course of the project two different micro spheres where 
investigated, Vaterite and liquid crystal droplets. Both can 
be readily synthesised in the lab and are will rotate at a 
variety of sizes (see \ref{sec:opt_torque}).

Vaterite is a polymorph of calcium carbonate that is rarely 
seen in nature due to its low stability \cite{KonopackaLyskawa2019}. 
All three polymorphs are inherently birefringent meaning 
that they can be rotated using circularly polarised light. 
However unlike its other polymorphs of calcite and aragonite, 
when synthesised vaterite will typically form small spherical 
particles making them ideal for optical trapping and rotation. 
Synthesis of Vaterite micro spheres requires fine control of 
the crystal growth process in order to maintain polymorphic 
stability. Though for the purposes of optical rotation the 
exact polymorph is not as important as its morphology as all 
3 polymorphs are inherently birefringent.

Vaterite samples where made by the first preparing equal amounts 
of $CaCl_2$ and $Na_2CO_3$ at a concentration of $0.33M$, at 
the same time a vial of $0.33M\ MgSO_4$ was prepared and set 
aside for later. Firsta small vial was filled with $1.5mL$ of 
$CaCl_2$ followed by $60\mu L$ and $90\mu L$ of $MgSO_4$ and 
$NaCO_3$ respectively, forming a seed solution. Next, a larger 
vial was filled with $5\ mL$, $1.5\ mL$, and $1\ mL$ of $CaCL_2$, 
$MgSO_4$, and $NaCO_3$ respectively followed by the seed solution. 
After 10 minutes of slow but continuous mixing a few drops of 
Agepon was added to halt the reaction, the solution was filtered 
and washed 3 times with distilled water before being suspended 
in water.

When trapped in circularly polarised light, the anisotropic 
crystal lattice allows spin angular momentum to be transferred 
to the Vaterite particle, resulting in a rotation about the 
beam axis. Because the particle scatters light anisotropically 
the signal detected by the QPD shows periodic fluctuations 
\cite{Yogesha2012}. In addition due to the change in polarisation 
there is also a periodic variation that appears at twice the 
rotational frequency \cite{Monteiro2018}. Therefore, the resulting 
power spectrum is not a Lorentzian but now also displays peaks 
that appear at integer multiples of the particles rotational 
frequency. 
\begin{figure}[h!]
	\centering
	\begin{subfigure}{0.4\linewidth}
		\includegraphics[width=\linewidth]{vaterite_sample.jpg}
		\subcaption{}
	\end{subfigure}
	\begin{subfigure}{0.55\linewidth}
		\includegraphics[width=\linewidth]{rotating_psd.png}
		\subcaption{}
	\end{subfigure}
	\caption{(a)Sample Vaterite sphere suspended in water and trapped by
		 circular polarised trap. (b) Collected power spectrum from 
		 rotating Vaterite, peaks in the power spectrum appear at integer 
		 multiples of the rotational frequency ($f_{rot} \approx 49.8$ Hz)}
	\label{fig:vaterite}
\end{figure}

As shown by Fig.~\ref{fig:vaterite}(b) the power spectra produced still 
demonstrates a Lorentzian curve but modified with these periodic peaks, 
while the Lorentzian can be loosely fitted to the end tail there exists 
no current model for describing the power spectra. The closest approximation
to this was conducted by \cite{Yogesha2012} where they describe the 
rotational motion of ellipsoidal polystyrene particles. The critical 
assumption being that the particle perfectly rotates in the $x-y$ plane. 
It has long been suspected that birefringent microspheres experience torques 
outside of the $x-y$ plane \cite{Volpe2023} making it very difficult to 
characterise the behaviour of rotating birefringent microspheres without a 
proper understanding of the full optical torque being applied to it.

\subsection{Liquid Crystal Rotors}
Liquid crystals are an intriguing example of materials with mixed phase
properties. Unlike typical solutes such as Glycine, a liquid crystal
can still maintain some degree of order between its individual molecules
while in the liquid state. This is due to the fact that liquid crystals
are constructed of ordered molecules that demonstrate a long range
ordering. There are three main types of liquid crystal transition methods:
Thermotropic crystals will transition to their liquid crystal phase when 
sufficiently heated. Lyotropic materials can undergo this transition due 
to changes in temperature and concentration. And lastly, Metallotropic 
materials - which are composed of both organic and inorganic molecules - 
change phase according to the ratio of organic to inorganic molecules present.
Liquid crystal rotors are rather simple in their production, 
4-Heptyl-4-biphenylcarbonitrile (7CB) was purchased from Sigma Aldrich
and a small amount was added to a vial of distilled water. The solution 
was then heated in a water bath to $25^\circ$ in order to transition the
solid crystal into its liquid crystal state. The solution can then be 
loaded onto a sample cover slip and the individual droplets visualised.
The molecules of 7CB will align with a strong electric field, and due to
the spherical droplet geometry the droplets are inherently birefringent. 
\begin{figure}[h!]
	\centering
	\includegraphics[width=0.5\linewidth]{LC_sample.png}
	\caption{Liquid crystal undergoing rotation due to the circularly polarised trap.} 
\end{figure}

The liquid crystal droplets had a much faster rotation rate than comparable 
Vaterite spheres, due to their higher degree of birefringence and the fact
that the droplets are far closer to perfect spheres making angular momentum
transfer more efficient.

>>>>

\chapter{Complex Langevin dynamics of spherical dimers}
\label{chapter:langevin_dynamics}
\section{Overview}
Often in calculation of optical forces and scattering the
choice of target particle is often either a single sphere 
or a large agglomeration of spheres. In optical tweezing 
research the target particle is not guaranteed to be a 
sphere, instead either being an asymmetrical shape, or being 
a collection of smaller spheres.

As an example, while working with dense colloidal suspensions, 
one often ends up trapping more than one sphere. Li and 
Arlt \cite{Li2008} studied the case of two microspheres 
trapped in a single OT and found that multiple trapped 
beads could be mistaken for a single trapped bead with an  
altered trap stiffness (see \ref{sec:harmonic_traps}). 
Theoretical studies on the case of two trapped microspheres 
by Xu~\textit{et~al.} \cite{Xu2005} employed a ray-optics 
based model to show that the two trapped beads are brought 
into physical contact with each other by optical forces 
and they also calculated the axial equilibrium positions 
of the two trapped beads as a function of their size. 
Experiments in \cite{Praveen2016} confirmed that the two 
trapped beads indeed experience different trap stiffnesses 
in the vicinity of the same potential well.

There are further discussions looking into the dynamics of 
a whole host of asymmetrically shaped particles (e.g. 
ellipsoids, cylinders, amorphous solids, disks, etc) 
\cite{Loudet2014, ShengHua2005, Chetana2022}, their results 
all showing that predicting the behaviour an arbitrary shaped 
particle comes with great difficulty due to the fact that the 
optical force is dependent on a greater number of variables 
such as orientation and size factors. While these have been 
noted as important factors, there has not been a concentrated
effort to discuss their impact over the course of a typical 
trajectory.

With the initial goal of the project being to induce 
nucleation events via a rotating sphere, the aim of this 
chapter is to - in a limited capacity simulate and investigate 
the influence of a second particle being bound to our target 
sphere. The choice of a dimer, instead of an amorphous solid 
that might better represent a growing crystalline solid, 
allows us to consider how the dynamics of the aggregate change 
by varying a single parameter, namely the size ratio of the 
two spheres. Additionally, attempting to simulate an amorphous 
aggregate is rather difficult as calculating the optical force 
and torque is computationally slow.

 In this chapter we will consider how the addition of a 
 second sphere changes the trapping dynamics by introducing 
 multiple equilibrium positions. In addition we demonstrate 
 that the angular dependency of the optical force allows for 
 multiple equilibrium orientations for a single particle. 
 Furthermore, we look at how dimers interact with circularly 
 polarised light, noting how our current understanding of 
 light and particle interactions does not accurately explain 
 the observed behaviour. We show that a dimer can undergo 
 gyroscopic precession by utilising this unique interaction 
 with circularly polarised light.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Simulation Details}
As a paradigmatic example, consider a dimer suspended in 
water ($n_p = 1.59, n_m = 1.33$) located at the focus of 
a Gaussian beam (more specifically a Laguerre-Gaussian 
beam of mode $[0,0]$, see \eqref{eq:incident}). The beam 
is focused by a objective with numerical aperture of 1.2.
The polarisation is defined using the Jones vector which 
describes the magnitude of the x and y components of the 
electric field (so the vector $[1.0, 0.0i]$ describes an 
electric field that only oscillates in the x direction). 
Unless stated otherwise the Jones vector is kept at $[1.0, 
0.0i]$.

As a paradigmatic example, we study a simple spherical 
dimer composed of polystyrene, while there is the 
possibility of considering absorption we choose to keep 
the refractive index fixed at 1.59. The ratio of the two 
sphere radii is given by $a_{I}/a_{II}$ where $a_{I}$ is 
kept at $1\ \mu m$ unless specified otherwise. For 
simulations we assume that the dimer is suspended in water 
at room temperature ($T = 298K,\ \eta = 1.0x10^{-3}Pa\ s$).

In order to compute the magnitude of the Brownian forces we
need to know the dimer's diffusion tensor \textbf{D}. Unlike
for spheres where the translational and rotational diffusion 
coefficients are equal along all three Cartesian axis', a
dimer's translational and rotational diffusion is dependent
on whether we consider the its long or short axis. We use
the results from \cite{Nir1973} to compute the dimensionless 
diffusion tensor and then scale it according to the radius of
the largest sphere ($a_I$). The dimer's centre of diffusion
is located a distance of $a_I\zeta$ from the point of contact
between the two spheres, the coefficient $\zeta$ is based on
the size ratio of the two spheres. We use the centre of 
diffusion as the target origin that is provided to \textit{mstm}

Where $U_3$ represents the orientation vector that passes
through the centres of both sphere's (see \ref{sec:sim} 
for a full breakdown), $U_1$, and $U_2$ are orthogonal to 
$U_3$ but have no physical representation for a dimer. We 
can represent $U_3$ by it's spherical coordinates ($\theta$ 
and $\phi$), where $\theta$ is the angle between the 
orientation vector and the direction of propagation and 
$\phi$ as the angle between the orientation vector and 
the x-axis. We define the 'standard' orientation as 
being aligned with the direction of beam propagation 
direction ($\theta=0^\circ$) - and therefore the 'inverted' 
orientation is defined when the dimer is orientated against 
the direction of beam orientation ($\theta=180^\circ$) - see
fig~\ref{fig:paradigmatic} for a visual of the laboratory and
particle frame.

Throughout this project we use the work of Vigilante \cite{Vigilante2020} 
to perform a systematic study of the dynamics demonstrated by 
asymmetric dimers in both plane and circularly polarised light. 
The simulations have two primary frames of reference that require 
explanation. Firstly, there is the laboratory frame which 
describes the coordinate system, we define the origin of our 
simulation as the focus of our trapping beam. While the choice in 
direction for the $x$ and $y$ axis are arbitrary the $z$ axis is 
specifically chosen so that it is aligned with the direction of 
beam propagation. 
\begin{figure}[h!]
	\centering
	\includegraphics[width=0.6\linewidth]{lab_frame.png}
	\caption{Ray scattering render of both the laboratory frame and 
		particle frame for a dimer and a Gaussian beam. The origin 
		of the laboratory frame is defined as the focus of the 
		trapping beam, with the positive $z$-axis being aligned with 
		the direction of beam propagation. The particle frame is 
		centred on the dimer's centre of mass with the primary axis 
		$U_z$ being aligned with the centres of the two spheres. Black 
		line is scaled to represent $1\ \mu m$.}
	\label{fig:lab_frame} 
\end{figure}

Secondly there is the particle frame which describes the orientation 
state of the particle, orientations are reported as quaternions which 
can be readily transformed into rotation matrices. The dimer's 
orientation is given by a rotation matrix in the form: 
\begin{align}
	\begin{pmatrix}
		U_{1,x} & U_{2,x} & U_{3,x} \\
		U_{1,y} & U_{2,y} & U_{3,y} \\
		U_{1,z} & U_{2,z} & U_{3,z}
	\end{pmatrix}
\end{align}
The matrix describes the direction of the particles 3 orientation 
axis $U_1$, $U_2$, and $U_3$. The latter is oriented so that it 
is collinear with the centres of the two spheres; the direction 
of vector $U_3$ is such that it goes from the centre of $a_{II}$ 
to the centre of $a_{I}$. The origin of the 3 axis' is set on the 
particle's centre of mass, this is also used to define it's 
displacement from the origin of the laboratory frame. The diffusion 
tensor of the dimer is more complex than a sphere's as we must 
consider the rotational effects experienced by the dimer. We use 
the analytical solution of Nir \& Acrivos \cite{Nir1973}, where 
they provided the solution coefficients depending on the dimer's 
size ratio ($a_I/a_{II}$). Rather than recalculate the coefficients 
for every possible configuration of dimer we use a spline fitting 
function to approximate the diffusion tensor.

We utilise a python library that utilises \textit{mstm} and
\textit{ott} to simulate the dynamical behaviour of spherical
aggregates. By specifying the optical properties of the target 
\textit{mstm} generates the targets T-matrix which can then be 
used by \textit{ott} to compute the optical forces generated by 
a focused beam. Using the Finite Differences approach we can solve
the Langevin Equation to take discrete time steps and update the 
dimer's position and orientation. A time step of $10 \mu s$ is 
used in line with previous works \cite{Vigilante2020}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Positional and Orientational dependence of Trapping forces}
\label{sec:eq_positions}
First, let us look at how the addition of a second sphere
changes how we go about looking for equilibrium positions
relative to the beam focus. Typically when trapping a target
particle we do not care where it is spacially in relation 
to the trap focus, we assume that it will end up in the same 
equilibrium position regardless. This assumption can not 
be taken when considering a dimer, or any aggregate of spheres.

If we wanted to start from first principles and determine 
the trap strength on our target particle the first step 
would be to locate the equilibrium position relative to 
the trap focus. For a single sphere it is easy to enough 
to understand that its centre of mass will be drawn to 
focal point of the laser due to gradient forces, once there 
the force is analogous to a harmonic spring with a fixed 
trap stiffness (see fig.\ref{fig:harmonic_trap}). Now, if 
we consider instead a dimer, we now have two spheres both 
being drawn to the focus along by the same gradient force; 
in addition the scattering force is significantly more complex 
due to both spheres scattering the electromagnetic fields. 
This mutual scattering between individual spheres is what 
makes simulating spherical aggregates far more difficult 
compared to a single sphere, and even harder still to 
predict the position where the dimer's centre of mass is 
in equilibrium.

Because the scattering force is only significant in the 
direction of beam propagation \cite{Capitanio2002} we 
can assume there will only be a single potential well 
lying at the centre of the beam. Along the beam axis, 
such an assumption is no longer valid due to fact that 
spherical aggregates experience inter-particle scattering. 
This is a key reason for using \textit{mstm} as it 
accounts for that behaviour. The methodology for computing 
optical forces has been covered extensively for a number 
of different trapping conditions \cite{RanhaNeves2019}. 
So it is relatively easy to compute the trapping force 
and determine the equilibrium position by finding the 
position that minimises the net optical force and the 
local gradient is negative ($\delta F/\delta x < 0$) - 
we can assume that for a dielectric sphere the optical 
torque is negligible. For a dimer (or any arbitrary 
spherical aggregate), we now must consider both its 
position and orientation and find where the net optical 
force and torque are minimised.

After computing its $T$-matrix via \textit{mstm} and 
supplying that to \textit{ott} we compute the optical 
force exerted by a $50\ mW$ laser via \eqref{eq:optical_force} 
in the axial direction while the dimer is in its 'standard' 
orientation. As expected we see a single point where 
the dimer will be in equilibrium, the linear fit in 
fig.~\ref{fig:paradigmatic}(a) indicates a that the 
force can be modelled as a harmonic potential close to 
the equilibrium position ($F_z\approx-\kappa_z z$). 
The second point where the axial force goes to 0 cannot 
be considered as equilibrium position as the positive 
gradient indicates that the trap is unstable unless 
Brownian motion is ignored.

We repeated the same calculation but now while the 
dimer is in its 'inverted' orientation, instead of 
a single point where the optical force is minimised 
we see that there are instead two separate equilibrium 
positions, one above the focus and one below the 
focus. In this particular example the two positions 
are far enough apart that both can be considered as 
separate harmonic traps.    
\begin{figure}[h!]
	\centering
	\begin{subfigure}{.65\linewidth}
		\includegraphics[width=\linewidth]{lam=2_theta=0.png}
		\caption{}
		\label{lam=2}
	\end{subfigure}\hfill % <-- "\hfill"
	\begin{subfigure}{.25\linewidth}
		\centering
		\raisebox{60pt}[0pt][0pt]{\makebox{}\includegraphics[width=0.3\linewidth, keepaspectratio]{theta=0.png}}
		\label{large over small}
	\end{subfigure}
	\medskip
	\begin{subfigure}{.65\linewidth}
		\includegraphics[width=\linewidth]{lam=2_theta=180.png}
		\caption{}
		\label{lam=2_inverted}
	\end{subfigure}\hfill % <-- "\hfill"
	\begin{subfigure}{.25\linewidth}
		\centering
		\raisebox{40pt}[0pt][0pt]{\makebox{}\includegraphics[width=0.3\linewidth, keepaspectratio]{theta=180.png}}
		\label{small_over_large}
	\end{subfigure}
	\caption{Plots of force vs displacement of the 
		centre of mass of the dimer (µm) for the 
		case of a dimer of size ratio 2. (a) is the 
		case where the dimer is in its' 'standard' 
		orientation, where the dimer is trapped at 
		$z = 1.71\ \mu m$. (b) is the case where the 
		dimer is in its' 'inverted' orientation, the 
		dimer is trapped at two positions: $z = 1.20\ 
		\mu m$ \& $z=-1.63\ \mu m$. On the left are 
		renders to visualise the dimer orientation 
		are shown below each plot. The black lines 
		on each force-curve is a linear fit with the 
		slope being reported as the trap stiffness in 
		the legend.}
	\label{fig:paradigmatic}
\end{figure}

This is different to a single sphere, namely in that a
single sphere should only have one equilibrium position
and that it is much closer to the trap focus ($\approx
0.05\mu\ m$ from the focus). This is also not similar 
to other works where random clusters of spheres have 
been shown to oscillate between stable points 
\cite{Vigilante2020}.

We can see that both equilibrium positions have 
comparable axial trap stiffness ($\kappa_z$), 
however the difference in the transverse trap 
stiffness ($\kappa_x$) is far more noticeable. 
The same dimer was trapped at each of the axial 
equilibrium positions and the transverse force was 
evaluated. While in all three cases the dimer can be 
trapped the linear range where that would typically 
associated with a stable trap is far narrower in the 
'standard' orientation compared to the 'inverted' cases. 
This highlights one of the challenges involved with 
studying asymmetric particles, even though its a 
simple enough process to trap them they maybe 
characterised very differently depending on their 
relative position and orientation towards the focus. 
This can have a significant impact on rheological 
studies - or attempting to probe any local property 
- as the variance in trap strength can result in 
large errors over repeated measurements. 
\begin{figure}[h!]
	\centering
	\includegraphics[width=\linewidth]{transverse_force.png}
	\caption{Plots of force vs displacement of the dimer's 
		centre of mass spheres, where a positive force 
		indicates the dimer is directed right on the x-axis, 
		and vice versa for a negative force. The same simulation 
		parameters are used here as in fig~\ref{fig:paradigmatic}
		(a) and (c). The blue curve representing the force 
		response for a dimer in its standard orientation, orange 
		being the inverted case, and green the same case but 
		placed below the focus.}
	\label{fig:transverse_force}
\end{figure}

\newpage
For completeness the harmonic traps were located for dimers
across a range of size ratios - from $a_{I}/a_{II} = 1$ to 
$a_{I}/a_{II}=10$ - while also recording the trap stiffness 
for each trap. The same simulation parameters are used here 
as for figures \ref{fig:paradigmatic} \& \ref{fig:transverse_force}. 
As shown in Fig.~\ref{fig:eq_positions} $a_{II}$ decrease 
the dimer begins to approximate a single homogenous sphere 
- at least in terms of location and trap strength. However, 
for intermediate sized dimers (between $a_{I}/a_{II} = 1.1$ 
to $a_{I}/a_{II}=4$), a second equilibrium position is found 
below the trapping focus. We should note that there are also
unstable traps that exist, these are also depicted in 
fig.~\ref{fig:eq_positions}. While the optical force
is zero at these positions the gradient $\delta F_z/\delta z$
is positive and so any Brownian motion would displace the 
dimer, either ejecting it from the optical trap or letting
it be drawn towards another harmonic trap. 
\begin{figure}[h!]
	\centering
	\begin{subfigure}{0.85\linewidth}
		\includegraphics[width=\linewidth]{Equillibrium_positions.png}
		\caption{}
		\label{eq_pos}
	\end{subfigure}\hfill % <-- "\hfill"
	\medskip
	\begin{subfigure}{0.85\linewidth}
		\includegraphics[width=\linewidth]{Equillibrium_positions_inverted.png}
		\caption{}
		\label{eq_pos_inverted}
	\end{subfigure}
	\caption{Equilibrium positions of optically trapped dimers 
		with varying size ratio, dashed lines represent unstable 
		traps whereas solid lines are for stable equilibrium 
		positions. (a) shows that dimers while in their 
		'standard' orientation will always have a single 
		equilibrium position. (b) shows that when the same dimer 
		is in its' 'inverted' orientation can be trapped in two 
		axial positions, one below the focus and one above the 
		focus.}
	\label{fig:eq_positions}
\end{figure}

Previous work using the ray-optics model have confirmed even 
in the case that two spheres begin separated the electric field 
will align the particles as such that they make contact and 
are trapped together about a single trapping position 
\cite{Xu2005}. Furthermore it has been shown through proper 
manipulation of the Gaussian or Bessel beam modes that any 
number of trapping potentials can be developed 
\cite{Shahabadi2020} for nanoparticles. This result however, 
is the first example of an orientation dependent trapping 
situation using only a $TEM_00$ beam. Typical experimental 
arrangements cannot determine much information on the axial 
position of a trapped particle relative to the trap focus; 
this result indicates not only that dimers can be trapped 
in multiple axial positions but also their trapping behaviour 
is heavily dependent on the axial position. As such it is 
necessary that positional information in the z-axis can be 
elucidated if multiple spheres are trapped simultaneously.

\newpage
\section{Non-trivial basins of attraction}
\label{sec:off-axis}
As mentioned previously in \ref{sec:opt_torque} the 
torque applied to an anisotropic scatterer can be 
broken down into two categories: torque generated 
from the field polarisation and torque generated 
by the gradient force of the electric field. The 
former is generally only applicable for objects 
that are birefringent whereas the latter is 
applicable to dimers due to their elongated shape.
Since both spheres are attracted towards the centre
of intensity via the gradient force a dimer will 
experience a restoring torque that keeps the it
in an equilibrium orientation. For any elongated 
particle the torque is minimised when the long 
axis is aligned either parallel or perpendicular
with the direction of the electric field. In this 
section we try to find equilibrium configurations 
(positions and orientations where both the optical
force and torque are minimised).
